Inducing Domain Specific Languages for Bayesian Program Learning

### 2018-02-01 paper.tex
minor edits
proposed the name Helmholtz hacker
better list function DSL table

\newcommand{\system}{EC2.0~}

@@

Additionally, for each proposed $\mathcal{D}$ we have to reestimate $\theta$.
Although this problem may seem 
very similar to estimating the parameters of a probabilistic context free grammar (PCFG: see~\cite{}),
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
$\mathcal{D}$
 is actually context-sensitive
due to the presence of variables in the programs and also due to the
polymorphic typing system.
In the Appendix we derive a tractable MAP estimator for $\theta$.

\begin{algorithm}[tb]
   \caption{DSL Induction Algorithm}
   \label{grammarInductionAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
     \WHILE {true}
     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
     \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
     \STATE $\mathcal{D}\gets\mathcal{D}'$
     \ENDWHILE
   \end{algorithmic}
\end{algorithm}

@@

@@

\begin{table}
\centering
\begin{tabular}{| l | l |}
  \hline
  \emph{name} & \emph{type} \\
  \hline
    empty & \texttt{tlist($\alpha$)} \\
    singleton & \texttt{arrow($\alpha$, tlist($\alpha$))} \\
    range & \texttt{int $\rightarrow$ tlist(int)} \\
    concat & \texttt{tlist($\alpha$) $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
    map & \texttt{($\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\beta$)} \\
    reduce & \texttt{($\beta$ $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ $\beta$ $\rightarrow$ tlist($\alpha$) $\rightarrow$ $\beta$} \\
    \hline
    true & \texttt{bool} \\
    not & \texttt{bool $\rightarrow$ bool} \\
    and & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
    or & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
    \hline
    0, $\ldots$, 9 & \texttt{int} \\
    + & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
    * & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
    negate & \texttt{int $\rightarrow$ int} \\
    mod & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
    eq? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
    gt? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
    is-prime & \texttt{int $\rightarrow$ bool} \\
    is-square & \texttt{int $\rightarrow$ bool} \\
    sort & \texttt{tlist(int) $\rightarrow$ tlist(int)} \\
    \hline
    sum & \texttt{tlist(int) $\rightarrow$ int} \\
    reverse & \texttt{tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
    all & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ tlist($\alpha$) $\rightarrow$ bool} \\
    any & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ tlist($\alpha$) $\rightarrow$ bool} \\
    index & \texttt{int $\rightarrow$ tlist($\alpha$) $\rightarrow$ $\alpha$} \\
    filter & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
    slice & \texttt{int $\rightarrow$ int $\rightarrow$ tlist($\alpha$) $\rightarrow$ tlist($\alpha$)} \\
  \hline
\end{tabular}
\caption{DSL for the domain of list function.}
\label{listdsl}
\end{table}

@@

\begin{algorithm}[tb]
   \caption{DSL Learner}
   \label{mainAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
     \STATE \textbf{Hyperparameters:} Frontier size $F$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, bottom-up recognition model $q(\cdot)$
     \STATE Initialize $\mathcal{D}_0\gets \mathcal{D}$, $\theta_0\gets \text{uniform}$, $q_0(\cdot ) = \theta_0$
     \FOR{$i=1$ {\bfseries to} $I$}
     \FOR{$x:\tau\in X$}
     \STATE  $\mathcal{F}_x\gets \{z| z\in \text{enumerate}(\mathcal{D}_{i - 1},q_{i - 1}(x),F)\cup\text{enumerate}(\mathcal{D}_{i - 1},\theta_{i - 1},F) \text{ if }\probability[x|z] > 0\}$
     \ENDFOR
     \STATE $\mathcal{D}_i,\theta_i\gets $induceGrammar$(\{\mathcal{F}_x\}_{x\in X})$
     \STATE Define $Q_x(z) \propto \begin{cases}
       \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
       0&x\not \in \mathcal{F}_x
     \end{cases}$
     \STATE $q_i\gets \argmin_q \sum_{x\in X}\text{KL}(Q_x(\cdot )||\probability[\cdot |\mathcal{D}_i,q(x)])$
      \ENDFOR
 \STATE \textbf{return} $\mathcal{D}^I,\theta^I,q^I$
\end{algorithmic}
\end{algorithm}

>>>>

\newcommand{\system}{\textsc{HelmholtzHacker}~}

@@

Additionally, for each proposed $\mathcal{D}$ we have to reestimate $\theta$.
Although this problem may seem 
very similar to estimating the parameters of a probabilistic context free grammar (PCFG: see~\cite{}),
for which we have effective approaches like the Inside/Outside algorithm~\cite{international2000derivation},
$\mathcal{D}$
 is actually context-sensitive
due to the presence of variables in the programs and also due to the
polymorphic typing system.
In the Appendix we derive a tractable MAP estimator for $\theta$.
\begin{algorithm}[tb]
   \caption{DSL Induction Algorithm}
   \label{grammarInductionAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Set of frontiers $\{\mathcal{F}_x\}$
     \STATE \textbf{Hyperparameters:} Pseudocounts $\alpha$, regularization parameter $\lambda$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$
     \STATE Define $L(\mathcal{D},\theta) =  \prod_x \sum_{z\in \mathcal{F}_x} \probability[z|\mathcal{D},\theta]$
     \STATE Define $\theta^*(\mathcal{D}) = \argmax_\theta \text{Dir}(\theta|\alpha) L(\mathcal{D},\theta)$
     \STATE Define $\text{score}(\mathcal{D}) = \log \probability[\mathcal{D}] + L(\mathcal{D},\theta^*) - ||\theta||_0$
     \STATE $\mathcal{D}\gets$ every primitive in $\{\mathcal{F}_x\}$
     \WHILE {true}
     \STATE N $\gets \{\mathcal{D}\cup \{s\} | x\in X, z\in \mathcal{F}_x, s\text{ subexpression of }z\}$
     \STATE $\mathcal{D}'\gets \argmax_{\mathcal{D}'\in N}\text{score}(\mathcal{D}') $
     \STATE \textbf{if }$\text{score}(\mathcal{D}') < \text{score}(\mathcal{D})$\textbf{ return }$\mathcal{D},\theta^*(\mathcal{D})$
     \STATE $\mathcal{D}\gets\mathcal{D}'$
     \ENDWHILE
   \end{algorithmic}
\end{algorithm}

@@

\noindent\textbf{Putting it all together.}
Algorithm~\ref{mainAlgorithm} describes how we combine the
program search, recognition model training, and DSL induction.

\begin{algorithm}[tb]
   \caption{The \system Algorithm}
   \label{mainAlgorithm}
   \begin{algorithmic}
     \STATE {\bfseries Input:} Initial DSL $\mathcal{D}$, set of tasks $X$, iterations $I$
     \STATE \textbf{Hyperparameters:} Maximum frontier size $F$
     \STATE \textbf{Output:} DSL $\mathcal{D}$, weight vector $\theta$, bottom-up recognition model $q(\cdot)$
     \STATE Initialize $\mathcal{D}_0\gets \mathcal{D}$, $\theta_0\gets \text{uniform}$, $q_0(\cdot ) = \theta_0$
     \FOR{$i=1$ {\bfseries to} $I$}
     \FOR{$x:\tau\in X$}
     \STATE  $\mathcal{F}_x\gets \{z| z\in \text{enumerate}(\mathcal{D}_{i - 1},q_{i - 1}(x),F)\cup\text{enumerate}(\mathcal{D}_{i - 1},\theta_{i - 1},F) \text{ if }\probability[x|z] > 0\}$
     \ENDFOR
     \STATE $\mathcal{D}_i,\theta_i\gets $induceDSL$(\{\mathcal{F}_x\}_{x\in X})$
     %% \STATE Define $Q_x(z) \propto \begin{cases}
     %%   \probability[x|z]\probability[z|\mathcal{D}_i,\theta_i]&x\in \mathcal{F}_x\\
     %%   0&x\not \in \mathcal{F}_x
     %% \end{cases}

     \STATE $q_i\gets \text{train recognition model to maximize }\mathcal{L}_{\text{RM}}$
      \ENDFOR
 \STATE \textbf{return} $\mathcal{D}^I,\theta^I,q^I$
\end{algorithmic}
\end{algorithm}

@@

\begin{table}
\centering
\begin{tabular}{| l | l |}
  \hline
  \emph{name} & \emph{type} \\
  \hline
    empty & \texttt{[$\alpha$]} \\
    singleton & \texttt{$\alpha$ $\rightarrow$ [$\alpha$]} \\
    range & \texttt{int $\rightarrow$ [int]} \\
    concat & \texttt{[$\alpha$] $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
    map & \texttt{($\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ [$\alpha$] $\rightarrow$ [$\beta$]} \\
    reduce & \texttt{($\beta$ $\rightarrow$ $\alpha$ $\rightarrow$ $\beta$) $\rightarrow$ $\beta$ $\rightarrow$ [$\alpha$] $\rightarrow$ $\beta$} \\
    \hline
    true & \texttt{bool} \\
    not & \texttt{bool $\rightarrow$ bool} \\
    and & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
    or & \texttt{bool $\rightarrow$ bool $\rightarrow$ bool} \\
    \hline
    0, $\ldots$, 9 & \texttt{int} \\
    + & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
    * & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
    negate & \texttt{int $\rightarrow$ int} \\
    mod & \texttt{int $\rightarrow$ int $\rightarrow$ int} \\
    eq? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
    gt? & \texttt{int $\rightarrow$ int $\rightarrow$ bool} \\
    is-prime & \texttt{int $\rightarrow$ bool} \\
    is-square & \texttt{int $\rightarrow$ bool} \\
    sort & \texttt{[int] $\rightarrow$ [int]} \\
    \hline
    sum & \texttt{[int] $\rightarrow$ int} \\
    reverse & \texttt{[$\alpha$] $\rightarrow$ [$\alpha$]} \\
    all & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool} \\
    any & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ bool} \\
    index & \texttt{int $\rightarrow$ [$\alpha$] $\rightarrow$ $\alpha$} \\
    filter & \texttt{($\alpha$ $\rightarrow$ bool) $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
    slice & \texttt{int $\rightarrow$ int $\rightarrow$ [$\alpha$] $\rightarrow$ [$\alpha$]} \\
  \hline
\end{tabular}
\caption{DSL for the domain of list function.}
\label{listdsl}
\end{table}

@@
